\textbf{Dynamic Hover Detection:\\Creating GUIs in Elm}

### 2013-12-15 paper/paper.tex
Revise section 4. Not done yet.

Having established the need for \texttt{hoverablesJoin} and implementing it, we
now turn to stated task of creating menus. A menu can be thought of as an
instance of a tree.

@@

Using \texttt{plainText : String -> Element}, we can take a menu specification
of type \texttt{Tree (Signal String)} and turn into a \texttt{Tree (Signal
Element}. From there, we can use \texttt{hoverablesJoin} to get the hover
information of each element and create a structure of type \texttt{Tree (Signal
Element, Signal Bool)} to represent our menus.  By recursing on the combinator
\texttt{combine : {[}Signal a{]} -\textgreater{} Signal {[}a{]}}, we are able to
turn this structure into a \texttt{Signal (Tree (Element, Bool))}. By changing
our internal structure into a signal of trees instead of a tree of signals, we
can do the rest of our work with pure functions that operate on the tree, rather
than on individual signals.  For example, our rendering function
\texttt{renderMenu : Tree (Element, Bool) -\textgreater{} Element} can simply be
lifted onto the menu structure.  Similarly to limiting the use of the IO monad
in Haskell code, it is advantageous to limit the use of signals in Elm code.

@@

Also note that the input to our menu is a \texttt{Tree (Signal String)}, but
not a \texttt{Signal (Tree String)}. We are able to map from
\texttt{Signal Element -\textgreater{} (Signal Element, Signal Bool)},
but not from \texttt{Element -\textgreater{} (Element, Bool)}.
Therefore, if we tried to turn a \texttt{Signal (Tree String)} into a
\texttt{Signal (Tree (Element, Bool))}, we would need to lift an
impossible hoverable function of type
\texttt{Element -\textgreater{} (Element, Bool)} on each element.
Therefore, each individual menu element can contain dynamic information,
but the menu structure must be static. We can, in practice, get around
this restriction by creating a tree that is larger than necessary and
filling the unused nodes with empty string, which our implementation
handles appropriately.

A possible solution for that issue would be to write a function of type
\texttt{Signal (Tree a) -\textgreater{} Tree (Signal a)}. Although the
\texttt{combine} combinator can turn a list of signals into a signal of
lists, there isn't an inverse to turn a signal of lists into a list of
signals. It is impossible to write the function
\texttt{split : Signal {[}a{]} -\textgreater{} {[}Signal a{]}} because
the size of the list in the input is dynamic, whereas the size of the
output list is static. Since the \texttt{Tree} data structure contains
lists, it is therefore impossible to write a function of type
\texttt{Signal (Tree a) -\textgreater{} Tree (Signal a)}.

@@

We present these techniques and analysis in the hopes they are of
service to the Elm and FRP communities. We hope this paper becomes
useful in Elm's goal of making GUIs simpler to implement and more robust
to use.

>>>>

Having established the need for \texttt{hoverablesJoin} and implementing it, we
now turn to stated task of creating menus.
\subsection{Menu Representations}

A menu specification, the abstract model of a menu, can be thought of as an
instance of a tree.

@@

The obvious type for a menu structure that may change is \texttt{Signal (Tree
String)}. It is easy enough to convert a value of this type to one of type
\texttt{Signal (Tree Element)}, but we cannot add hover detection to this tree.
We are able to map from \texttt{Signal Element -\textgreater{} (Signal Element,
Signal Bool)}, but not (in any useful way) from \texttt{Element -\textgreater{}
(Element, Bool)}. That is, because the tree is already a signal, we cannot
create signals inside it (without creating signals of signals). Hover
information requires a (distinct) signal at each item in the tree.

Instead of a signal of trees, we therefore use a tree of signals. Each
individual menu element can contain dynamic information, but the menu structure
must be static. We can, in practice, get around this restriction by creating a
tree that is larger than necessary and filling the unused nodes with empty
string, which our implementation handles appropriately.

Note that we can (and will) convert a tree of signals into a signal of trees, it
is impossible to go the other way. When converting many signals into one, the
information on which signal changed is lost. The combinator \texttt{combine :
    {[}Signal a{]} -\textgreater{} Signal {[}a{]}} does this, and will be used
    below. It has no inverse because such a function would need to create
    information: which signal in the list updated? What if the list
    changes length? What if there is a repeat event, where the list does not
    change, which signals should propagate repeats? Similar arguments show why
    the mentally convenient type of \texttt{lift hoverable}, which turned a
    signal of pairs into a pair of signals, cannot actually be obtained.

\subsection{Menu Rendering}

Therefore, a menu is specified by a value of type \texttt{Tree (Signal String)}.
We can lift and map the combinator \texttt{plainText : String -> Element} to
obtain from a menu specification a value of type \texttt{Tree (Signal Element)}.
From there, we can use \texttt{hoverablesJoin} to get the hover information of
each Element and create a value of type \texttt{Tree (Signal Element, Signal
Bool)} to represent our menus. \emph{Now} we convert (by structural recursion)
our tree of signals to a signal of trees, a value of type \texttt{Signal (Tree
(Element, Bool))}.  We do the rest of our work with pure functions lifted to
operate on the tree, for example, our rendering function \texttt{renderMenu :
    Tree (Element, Bool) -\textgreater{} Element}. Similarly to limiting the use
    of the IO monad in Haskell code, it is advantageous to limit the use of
    signals in Elm code.

@@

@@

We present these techniques and analysis in the hopes they are of
service to the Elm and FRP communities. There are a number of obvious paths that
dead-end, and non-obvious paths that are fruitful. We hope this paper becomes
useful in Elm's goal of making GUIs simpler to implement and more robust
to use.
